---
title: "Mexico_city_Toronto_comparison"
author: "James Johnson"
date: "February 1, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(lubridate)
library(readxl)
library(gt)

MC_data <- readRDS("data/MC_data2013-2017.Rds")
MC.poll.labs <- c(bquote("NO" [2] ~ "(ppb)"), bquote("PM" [2.5] ~ "("*mu * g/m^3*")"), bquote("O" [3] ~ "(ppb)"))

```

## Assignment 2

- Identify a highly polluted city in the world (or small geographic region)
- Justify your selection of being highly polluted based on available data and standards/resources
-Describe condisitons for several air pollutants
 - levesl, seasonalily, responsible sources, public awareness, current efforts the responsible government is taking to improve air quality.
 - Compare levels in that highly polluted place with Toronto for PM2.5, NO2, and O3
 
## Mexico City
 
### Topography

Mexico City is in an elevated basin surrounded on three sides by mountain ridges on to the east, south and west. There is a broad opening to the north and a narrow gap to the south-southwest. (Harvard) Two major volcanoes are located in the southeast. Within the basin, the mountains and frequent thermal inversions trap pollutants within the basin. The high altitude also makes combustion sources less efficient.

```{r mc_map}

```

 ### Air quality
 
```{r}

daily_MC <- MC_data %>%
  mutate(year = year(date),
         trimes = floor_date(date, unit = "3 months"),
         month = floor_date(date,unit = "1 month"),
         day = floor_date(date, unit = "1 day"),
         month_num = month(date)) %>% 
  mutate(season = ifelse(month_num %in% c(11,12,1,2),"Cool",ifelse(month_num %in% c(3,4,5),"Warm","Rainy"))) %>%
  mutate(season = fct_inorder(season)) %>% 
  filter(station == "HGM") %>% 
  gather(pollutant, value, NO2:SO2) %>%
  filter(value > 0) %>% 
  group_by(season, hour, pollutant) %>% 
  summarise(mean_hourly =mean(value))

daily_MC %>% 
  ggplot(aes(hour,mean_hourly, col = pollutant)) +
  geom_line() +
  facet_wrap(~season) +
  theme_light() +
  labs (title = "2013-2017 pollution at Hospital General Mexico (HGM)\naveraged by hour",
        x = "Hour of day", y = "pollutant")


day_Mc <-  MC_data %>%
  mutate(year = year(date),
         trimes = floor_date(date, unit = "3 months"),
         month = floor_date(date,unit = "1 month"),
         day = floor_date(date, unit = "1 day")) %>% 
  filter(station == "XAL"|station == "COY"|station == "BJU"|station == "HGM") %>%
  filter(PM25 >0) %>% 
  group_by(station, year, trimes, day) %>% 
  summarize(mean = mean(PM25,na.rm = T),quint_75 = quantile(PM25, 0.75,na.rm = T),quint_98 = quantile(PM25, 0.98,na.rm = T)) %>% 
  ungroup()

days_above <- day_Mc %>%
  mutate(above_25 = mean > 25,
         above_45 = mean > 45) %>% 
  group_by(station, year) %>% 
  summarize(above_WHO = sum(above_25), above_NOM = sum(above_45)) %>% 
  ungroup() %>% 
  gather(standard,days_above, above_WHO:above_NOM)

ggplot(days_above,aes(factor(year),days_above, fill = station)) + geom_bar(stat = "identity", position = "dodge") + facet_wrap(~standard) + theme_light() + labs(y = "Days above standard", x = "Air quality station")

Mc_24_hr <- day_Mc %>% 
  group_by(station, year) %>%
  summarize(quant_75 = quantile(mean, 0.75)) %>% 
  spread(station,quant_75) %>% 
  gt() 
Mc_24_hr

Mc_annual <- day_Mc %>% 
  group_by(station, year,trimes) %>%
  summarize(mean_quint = quantile(mean,0.75, na.rm = T)) %>% 
  ungroup() %>% 
  group_by(station,year) %>% 
  summarize(annual_mean = mean(mean_quint, na.rm = T)) %>% 
  spread(station, annual_mean) %>% 
  gt()
Mc_annual
ggplot(day_Mc, aes(day,quint_75,col = station)) + geom_line()
  
  

```
 
 


```{r}



monthly <- MC_data %>%
  mutate(year = year(date),
         month = floor_date(date,unit = "1 month"))%>% 
  filter(station == "XAL"|station == "COY"|station == "BJU"|station == "HGM") %>%
  gather(pollutant,value, c(NO2:O3)) %>% 
  filter(value > 0) %>% 
  group_by(year,month,pollutant,station) %>% 
  summarize(mean = mean(value)) %>% 
  ungroup %>% 
  mutate(month2 = as.Date(paste0("2013-", month,"-01"),"%Y-%m-%d")) %>% 
  mutate(pollutant = factor(pollutant, levels = c("NO2","PM25","O3"), labels = MC.poll.labs ))


ggplot(monthly, aes(month,mean, col = station)) +
  geom_path(alpha = 0.8) +
  geom_point(size = 1, alpha = 0.8) +
  facet_wrap(~pollutant,nrow = 3, labeller = label_parsed) +
  scale_x_date(date_breaks = "3 months",date_labels = "%y %b") +
  theme_light() +
  theme(axis.text.x = element_text(angle = 45))+
  labs(title = "Mexico City average monthly pollutant concentrations\n2013-2017", y = "Pollutant concentration", x = "Date (yy - mmm)")

MC_data %>% 
   filter(!is.na(PM25)) %>%
  filter(PM25>0) %>% 
   count(station) 



```
 
